---
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
  output: default
title: "Indiana Bloomington DSE R Notebook"
---
# Time & Date in R (1)
## Time and date in R

There are three built-in date and time data classes:
  - **Date**: Designed for data with dates only, no times;
  - **POSIXct**: Used when both date and time are present, *"ct" stands for calendar time*;
  - **POSIXlt**: Used when both date and time are present; enables easy extraction of specific components from the date/time data. *"lt" stands for local time.*
  
#### Date

  - `as.Date()` handles dates without times. 
    - Default format is **"yyyy-mm-dd"** and **"yyyy/mm/dd"**


```{r}
# create a date variable:
dt1 <- as.Date("2020-02-14")
dt1

```
  - Aside from the **default format**, the other date formats need to be specified when creating dates. The `as.Date()` function allows different input formats through its format argument.
  
  
```{r}
dt3 <- as.Date("01/08/2020", format = "%m/ %d/ %Y")
dt3

dt4 <- as.Date("February 29, 2020", format = "%B %d, %Y")
dt4

dt5 <- as.Date("20Jan20", format = "%d %b %y")
dt5
```

| Symbol |Meaning| Example |
|:-------|:-------:|:-------:|
|**%d**  |day as a number (01-31)|01-31|
|**%a**, **%A**|abbreviated weekday, unabbreviated weekday|Mon, Monday|
|**%m** |month (01-12)|01-12|
|**%b**, **%B**|abbreviated month, unabbreviated month|Jan, January|
|**%y**, **%Y**|2-digit year, 4-digit year|20, 2020|

***

The R library *chron* handles dates and times, but time zones are not accounted for.

- Dates are stored internally as the number of days or seconds from January 1, 1970. Those dates after Jan 1, 1970 are positive, while dates prior are negative. You can use `as.numeric()` to convert a Date to a number.

```{r}
n1 <- as.numeric(as.Date("20Mar20", format = "%d %b %y"))
n2 <- as.numeric(as.Date("02/20/2020", format = "%m/ %d/ %Y"))

n1
n2
n1 - n2 ## days difference between 03/20/20 and 02/20/2020
```

- Calculating dates in different units:
```{r}
dtSix <- as.Date("02/28/2020", format = "%m/ %d/ %Y")
dtSev <- as.Date("02/28/2019", format = "%m/ %d/ %Y")
difftime(dtSix, dtSev, units = "days")
## Time difference of 365 days
difftime(dtSix, dtSev, units = "weeks")
## Time difference of 52.14286 weeks
```

- Alternative units of time can be chosen:
  - units = `auto`, `secs`, `mins`, `hours`, `days`, `weeks`
  
- Components of dates can be extracted with the following functions:
  - `weekdays()`, `months()`, `quarters()`
  
```{r}
# Day of the week of each Christmas since 2016:

Cdays <- c(C2016 = as.Date("2016-12-25"), C2017 = as.Date("2017-12-25"), c(C2018 = as.Date("2018-12-25"), C2019 = as.Date("2019-12-25"), c(C2020 = as.Date("2020-12-25"))))

weekdays(Cdays)
```

