---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---
## Simulation(1)
  - Some common simulation functions:

  - Create a sample dataset with a class of 100 students, including student id, age, gender,     and grade variables.

```{r}
# generate fake students

students = data.frame(id = 1:100,
                      age = as.integer(rnorm(100, mean = 25, sd = 6)),
                      gender = sample(c("Male", "Female"), 100, replace = TRUE),
                      grade = sample(LETTERS[1:5], 100, replace = TRUE)
                      )
students
```
#### rnorm function
`rnorm()` generates a random normal variable. 
  - takes 3 arguments;
  - (num-of-elements-to-generate, mean, standard-deviation of normal dist)
  
#### runif function
`runif()` generates uniform random variates.
  - Takes 3 arguments:
    - (num-elements-to-gen, max-value, min-value)
    
```{r}
runif(10, 5, 15)
```

#### sample function
**Sampling** is the process of choosing a subset of a set of observations.
  - With replacement - a population element can be chosen more than once;
  - Without replacement - a population element can be selected only once.

  - 3 arguments: (population-vector, sample-`size`, `replace`)
    - `size` - size of the desired sample;
    - `replace` - determines if sampling is with or without replacement.
```{r}
x = 1:10

sample(x) # randomly sample population vector x, and generate sample vector of same length as vector x

sample(x) # sample again to get a different order

sample(x, size = 5) # generates a sample of 5 elements

sample(x, size = 5) # the elements chosen as well as their order will be different

```

`sample()` will generate a random permutation of vector `x`, containing elements from 1 to 10.

```{r}
sample(x, size = 15)
```
```{r}
sample(x, size = 15, replace = TRUE)
```

- Let's toss a coin 10 times to generate a vector `v` that contains `Heads` or `Tails` for each toss result. Need to sample with replacement with `replace = TRUE`.

```{r}
# sample of 10 consecutive coin tosses
sample(c("Heads", "Tails"), size = 10, replace = TRUE)
```

```{r}
# creating sample genders for 100 students
sample(c("Male", "Female"), 100, replace = T)
```

```{r}
# creating sample grades for 100 students
sample(LETTERS[1:5], 100, replace = T)
```

```{r}
# generate data of students in a class

students = data.frame(id = 1:100,
                      age = as.integer(rnorm(100, mean = 25, sd = 6)),
                      pronoun = sample(c("He/Him", "She/Her", "Ze/Zed"), 100, replace = TRUE),
                      grade = sample(LETTERS[1:5], 100, replace = TRUE)
                      )
students
```

## Simulation (2)
  ### Generating Missing Values
  
`NA` in R represents **missing values**.

To simulate `NA` in test data, use the `is.na()` function. 

```{r}
a = c(1:10)

is.na(a)
```

```{r}
a[4] = NA
a[7] = NA

is.na(a)
```

Another way, perhaps more efficient:
```{r}
a = c(1:10)

is.na(a) = c(4,7)

is.na(a)
```

```{r}
# sample function to randomly select 10 indices out of 1:100
# then is.na will assign NA to those locations:

is.na(students$age) = sample(1:100, 10)

is.na(students$age)
```

```{r}
str(students)

summary(students)
```

  ### Generating sequences
  
- Use `seq()` to make sequences of numbers flexibly. 
  - 3 arguments:
    - (starting-number, ending-number, `by` specifies an increment)
    
```{r}
# sequence 1 to 20 by increment of .5

seq(1, 20, by = 0.5)
```

- The function `rep()` repeats values. Makes a vector of repeated values.
```{r}
# create 10 zeroes:
rep(0, times = 10)

# repeats c(1,2) 10 times:
rep(c(1,2), times = 10)

# repeats first value 10 times and then second value another 10 times:
rep(c(1,2), each = 10)

# repeats 1111122222 five times
rep(c(1,2), times = 5, each = 5)
```

##### Questions:
  1. What will `rep(c(1,2), each = 5, times = 5)` give?

```{r}
rep(c(1,2), each = 5, times = 5)
```

