---
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
  output: default
title: "Indiana Bloomington DSE R Notebook"
---
# Time & Date in R (1)
## Time and date in R

There are three built-in date and time data classes:
  - **Date**: Designed for data with dates only, no times;
  - **POSIXct**: Used when both date and time are present, *"ct" stands for calendar time*;
  - **POSIXlt**: Used when both date and time are present; enables easy extraction of specific components from the date/time data. *"lt" stands for local time.*
  
#### Date

  - `as.Date()` handles dates without times. 
    - Default format is **"yyyy-mm-dd"** and **"yyyy/mm/dd"**


```{r}
# create a date variable:
dt1 <- as.Date("2020-02-14")
dt1

```
  - Aside from the **default format**, the other date formats need to be specified when creating dates. The `as.Date()` function allows different input formats through its format argument.
  
  
```{r}
dt3 <- as.Date("01/08/2020", format = "%m/ %d/ %Y")
dt3

dt4 <- as.Date("February 29, 2020", format = "%B %d, %Y")
dt4

dt5 <- as.Date("20Jan20", format = "%d %b %y")
dt5
```

| Symbol |Meaning| Example |
|:-------|:-------:|:-------:|
|**%d**  |day as a number (01-31)|01-31|
|**%a**, **%A**|abbreviated weekday, unabbreviated weekday|Mon, Monday|
|**%m** |month (01-12)|01-12|
|**%b**, **%B**|abbreviated month, unabbreviated month|Jan, January|
|**%y**, **%Y**|2-digit year, 4-digit year|20, 2020|


