---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---
## Simulation(1)
  - Some common simulation functions:

  - Create a sample dataset with a class of 100 students, including student id, age, gender,     and grade variables.

```{r}
# generate fake students

students = data.frame(id = 1:100,
                      age = as.integer(rnorm(100, mean = 25, sd = 6)),
                      gender = sample(c("Male", "Female"), 100, replace = TRUE),
                      grade = sample(LETTERS[1:5], 100, replace = TRUE)
                      )
students
```
#### rnorm function
`rnorm()` generates a random normal variable. 
  - takes 3 arguments;
  - (num-of-elements-to-generate, mean, standard-deviation of normal dist)
  
#### runif function
`runif()` generates uniform random variates.
  - Takes 3 arguments:
    - (num-elements-to-gen, max-value, min-value)
    
```{r}
runif(10, 5, 15)
```

#### sample function
**Sampling** is the process of choosing a subset of a set of observations.
  - With replacement - a population element can be chosen more than once;
  - Without replacement - a population element can be selected only once.

  - 3 arguments: (population-vector, sample-`size`, `replace`)
    - `size` - size of the desired sample;
    - `replace` - determines if sampling is with or without replacement.
```{r}
x = 1:10

sample(x) # randomly sample population vector x, and generate sample vector of same length as vector x

sample(x) # sample again to get a different order

sample(x, size = 5) # generates a sample of 5 elements

sample(x, size = 5) # the elements chosen as well as their order will be different

```

`sample()` will generate a random permutation of vector `x`, containing elements from 1 to 10.

```{r}
sample(x, size = 15)
```

