---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Module 3: Operators

### 4 Types of Operators in R:
  - Arithmetic
  - Logical
  - Assignment
  - Relational

## Arithmetic Operators

```{r}
x = 5
y = 2

x + y # addition

x - y # subtraction

x * y # multiplication

x / y # division

x ^ y # exponent

x %% y # modulus - remainder after division

x %/% y # integer division

```

## Relational Operators

  - Used to compare the values of two variables.
  - Always returns a boolean.
  
```{r}
x = 5
y = 4

x > y # greater than

x >= y # greater than or equal to

x < y # less than

x <= y # less than or equal to

x == y # equal to

x != y # not equal
```

  - With two vectors or matrices as operands, R applies the operator *elementwise*.
  
```{r}
x = c(5, 3, 6, 8, 9, 1)
y = c(1, 3, 4, 2, 6, 0)

x + y

x > y
```

  - If one operand is shorter in length than the other, the elements in the shorter one are recycled cyclically to match the length of the longer one.
  
```{r}
a = c(1, 3, 4)
b = c(1, 4, 6, 8)

a + x

b + x
```

## Logical Operators

  - Boolean operations

```{r}
x = c(T, F, 4, -1, 0) # zero evalutes to FALSE; any non-zero Numeric is TRUE
y = c(F, T, T, T, T)

x | y # element-wise logical OR

x || y # logical OR

x & y # element-wise Logical AND

x && y # logical AND

!x # logical NOT; a unary operator (takes only one operand)
```

  - `&` and `|` act element-wise on vectors; they recycle the shorter vector.
  - `&&` and `||` examine only the first element of the two operands and return a length 1 logical vector.
  
  - Create vector `marbles` that contains marbles in 55 containers.
    - `marbles[marbles > 165]` will find those containers with more than 165 marbles.
    - a logical vector inside the indexing operator, `[]`, we only get those where the logical vector has `TRUE` values.
  
```{r}
marbles = c(a = 150, b = 173, c = 184, d = 192, e = 163)

marbles > 165

marbles[marbles > 165]
```

## Assignment Operators

  - `->` is the "leftward assignment" operator, used in the same way as `<-` .
  - `<<-` and `->>` are used to modify R environments.

