---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Control Structures III

## break and next statement

  - `break` is used to stop a loop iteration completely, and continue on to the next command after the loop.
  
  - Loop is breaked after `i` becomes greater than `10`.

```{r}
x = 1:15
for(i in x){
  if(i > 10)
    break
  print(i)
}

```

  - `next` is used to skip the current iteration of the loop, and not terminate the loop completely, and continue to the next iteration.

```{r}
x = 1:5

for(i in x){
  if(i == 2){
    next
  }
  print(i)
}
```

## repeat Loop
  - iterate over a block of code multiple times. No stop condition test.
  - Instead, an explicit condition is put in the body of the loop, and `break` is used to exit it.
  - Failure to do so will result in an **infinite loop**.
  
```{r}
x = 0
repeat{
  print(x)
  x = x + 1
  if(x > 10)
    break
}

```

  - This loop is basically the same as below, with fewer lines of code:
```{r}
x = 0
while(x <= 10){
  print(x)
  x = x + 1
}
```
  
  