---
title: "Indiana Bloomington DSE R Notebook"
output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
   
---

## Return value of a function

  - `return()` function returns a value in R.

```{r}
pow <- function(x = 2, y = 2) {
  # function to return x raised to the power of y
  
  return(x^y)
}

pow(2,3)

```

  
```{r}
var = pow(2,3)
var
```

  - In the absence of *explicit returns* from a function, the **value of the last evaluated expression** is returned.
  
```{r}
pow <- function(x = 2, y = 2) {
  x^y
}

var = pow(2, 3)
var
```

  - To return multiple values, use a **list**.

```{r}
marble <- function(color, size) {
  my_list <- list(color = color, size = size)
  return(my_list)
}

x = marble("red", 10)
x
```

  - To avoid confusion between the names of list elements and names of parameters, it is best practice to give the names of **list elements** in *quotes*.
  
```{r}
marble <- function(color, size) {
  my_list <- list("color" = color, "size" = size)
  return(my_list)
  
}

x = marble("red", 10)
x
```
***
#### Exercises

  1. Write a function `check()` which checks whether a given integer is positive, negative, or zero. The function should take one parameter x and print the output.
  
```{r}
check <- function(num) {
  if (num > 0) {
    print("Positive")
  } else if (num < 0) {
    print("Negative")
  } else {
    print("Zero")
  }
}

check(3)
check(-3)
check(0)
```
  
  2. Write a function that takes 2 parameters, `rows` and `cols`, and returns a matrix with its elements assigned to the value you get from multiplying the row and column of that element.

  
```{r}
maketrix <- function(rows, cols) {
  l <- list()
  ct <- 1
  r <- c(1:rows)
  c <- c(1:cols)
  for (i in c) {
    for (j in r) {
      l[[ct]] <- c(i*j)
      ct <- ct + 1
    }
  }
  m <- matrix(l, nrow=rows, ncol=cols)
  return(m)
}

maketrix(5, 4)
```

## Are Functions also objects in R?

  - Functions are, in fact, **first class objects** .
  
### Environment

  - An R environment can be thought of as a *collection of objects*. Environments are created when the **R interpreter** starts. 
  - `R_GlobalEnv` : the top-level environment at the R command prompt.
    - Accessed using `.GlobalEnv` in R scripts.
  - `environment()` function gets the current environment.
    - Prints out `<environment:R_GlobalEnv>`

## Three components of functions

  1. **formal argument list**: `formals()` - list of arguments that controls how you can call the function.
  2. **body**: `body()` - the code inside the function.
  3. **environment**: `environment()` - environment in which the function is created.
  
  - `formals()` lists the formal arguments of a function.
  - `body()` - prints the code inside the function.
  
  - The environment of a function is the one that was active at the time the function was created.
  
```{r}
pow <- function(x = 2, y = 2) {
  return(x^y)
}

formals(pow)

body(pow)

environment(pow)
```

  - **Exception** 
    > **Primitive functions**, like `sum()`, call **C** code directly with `.Primitive()` and contain no **R** code. `formals()`, `body()`, and `environment()` return `NULL` for such functions.
    
```{r}
body(sum)
formals(sum)
environment(sum)
```


  - `is.primitive()` function checks whether a function is primitive or not.
  
```{r}
is.primitive(pow)

is.primitive(sum)
```
  
## Function Application

  - Apply a function to each element in a data structure:
    - for loop, perhaps;
    - `sapply` function:
    
      - `sapply` takes a vector or list as input, a function to apply element-wise along the data structure, and arguments for the function.
      
```{r}
vec = 1:5

sapply(vec,pow,y=2) # square each element

```
    
    