---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Functions I

  - Functions are self-contained blocks of code written to do a specific task.
  
## Writing Functions

  - A sample function to calculate exponents:

```{r}
pow <- function(x, y) {
  print(x^y)
}

pow(2, 3)
pow(4, 2)

```

## Function Arguments

  - Arguments in a function declaration are known as **formal arguments**. Those written into a function call directly are **actual arguments** .
  - Get all of the formal arguments of a function with `formals()`.
  
```{r}
formals(pow)
```

  - Pass `all.names = TRUE` to `ls()` to see the hidden variables in an R environment.
  - `stringsAsFactors` in the `data.frame()` function will prevent strings from converting to factors by default.
  - The `matrix()` function can use `byrow` argument to fill a matrix by rows.
  - **Positional order** is used to match formal arguments to actual arguments in a function call.
  - **Named arguments** : positional order does not matter. The formal argument names are used to match to actual arguments directly.
  
```{r}
pow(x = 2, y = 3)
pow(y = 3, x = 2)
```

  - In a call using both named and unnamed arguments, the named arguments are matched first, and then the unnamed arguments are matched in positional order.
  
```{r}
pow(2, y = 3)
pow(x = 2, 3)
```

  - **Default parameters** can be defined in the function declaration by giving a value to the argument directly.
  
```{r}
pow <- function(x = 2, y = 2) {
  print(x^y)
}

pow()
pow(2)
pow(, 3)

```

