---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Control Structures

## if, else Statements

  - R's `if` statement only takes a vector of length one.
  - It is best to use vectors of length one in the `test condition`.
  - Use `ifelse` to test a condition on vectors of length greater than one.

```{r}
x = 10

if(x > 0){
  print("x is Positive")
} else if(x < 0){
  print("x is Negative")
} else {
  print("x is zero")
}

```

## ifelse Function

  - **Vectorization** of a function is to take a vector as input, and alike return a vector. Vectorized functions is much faster than looping directly over the elements of a vector.
  - The *vector equivalent* of `if`, `else` statements is the `ifelse()` function.
  
  - The `test expression` must be a logical vector (or somehow coercible). `ifelse()` returns a vector of the same length as the `test expression`. 


```{r}
x = c(5, 0, 2, 3)
ifelse(x %% 2 == 0, "even", "odd")
```

