---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Control Structures II

## for Loop

  - To repeat a block of code a specified number of times.

  - The `sequence_vector` is a vector; the `loop_variable` takes on the value of each element during the loop.
  - `statement` gets evaluated.
```{r}
# Count the number of even elements in x

x = c(1, 2, 4, 5, 9, 7, 8)

count = 0
for(i in x){
  if(i %% 2 == 0)
    count = count+1
}

count

```
  - Nest a `for` loop inside another. Here, create a 10 x 10 matrix, assigning the value of each elementin the matrix the product of its row and column.
  
```{r}
a = matrix(nrow=10, ncol=10) # creates a 10 x 10 matrix

for(i in 1:dim(a)[1]) # one for each row
{
  for(j in 1:dim(a)[2]) # for each column
  {
    a[i, j] = i*j # assign as value the product of row and column index
  }
}

a
```

## while Loop

  - `while` loop is run until a condition is met.
  
```{r}
i = 10
while(i > 0){
  print(i)
  i = i - 1
}
```
  
  