---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Factors

  > Factor is a data structure used for fields that take on only a **predefined**, finite number of values, i.e. **categorical data**. 
  
## Create a factor

  - `factor()` to create | `levels()` to check levels
  
```{r}
blood_type = factor(c("A", "B", "AB", "O", "O", "B", "AB"))
blood_type

levels(blood_type)
```
## Properties of Factor

  - `typeof()` on a factor yields *integer*. `str()` shows the levels in a character vector, with each element stored as an integer index.
  
```{r}
typeof(blood_type)
str(blood_type)
```

## Modify a Factor

  - Similar to vectors for accessing and modifying elements.
  - A factor can take on values that are listed in its `levels` atrribute.
  
```{r}
x = factor(c("a", "b", "a", "b"))
x

```

```{r}
x[5] = "b"
x
```

```{r}
x[6] = "c" # gives error, as "c" is not a level
```

```{r}
# update the levels first:
levels(x) =c(levels(x), "c")
x[6] = "c"
x
```

