---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

## Modifying a Vector

- Modify in R using the **assignment operator**, `<-` or `=` .



```{r}
v = c(1:6)
v[2] = 12 # modify second element
v[1:3] =5 # modify elements from 1 to 3
v = v[1:4] # truncate vector to first 4 elements
v = c(v, 5, 6, 7)
v
```

Increase the length of a vector by **assigning** another element to the vector:

```{r}
v = 1:3
v[4] = 4
v
```

```{r}
v = 1:3
v[5] = 5
v[4]
```

```{r}
v = 1:5
length(v)

length(v) = 7
v
```

## Delete a Vector

Delete a vector by **assiging it the value `NULL`**, or use the `rm()` function. The latter will completely remove an object from the workspace.

```{r}
x = c(1:4)
x = NULL # assigns NULL to vector x
x # returns NULL

rm(x) # completely remove x
# x # Should error since it has been removed
```

## `str()` function

`str()` stands for **structure**. "What is this object?" -- gives contents of an object in abbreviated form.

```{r}
v = c(1:5)
str(v)
```

```{r}
x = c(1, 2, 3, 4, 5)
str(x)
```

## Vector Manipulation

```{r}
v = c(3,4,6,7,8)
x = 1:5
v + x

v - x

v/x

v*x

```

```{r}
v = c(2,3,5,6,7,9)
x = 1:2

v+x

v-x

v/x

v*x

```

## Coercion
- Combining a character and an integer yields a character, for example:

```{r}
v = c('a', 1, 3)
v
typeof(v)
```

- Logical vector coercion: **TRUE becomes 1**; **FALSE becomes 0**. Use `as.numeric()` to explicitly convert from logical to numeric.

```{r}
v = c(5, T, F)
v
x = c(T,F,F,T,F)
as.numeric(x) # convert the logical vector x to numeric

```

## Vectors of length 1

- Variable assignment `v = 3` does not create a numeric variable with value 3 as in other languages. It creates a vector object of length 1 and type numeric. *One cannot create a variable of primitive data types in R.*

#### Questions:

1. Create a vector v containing the integers from 1 to 10. What will the following subset operations give? v[] and v[0]

```{r}
v = 1:10
v[]
v[0]
```


2. Define a vector v = c(1, 4, 5, 7). It has length 4. What will happen if we try to access an element which is out of index, like v[5] or v[10]? 

```{r}
v = c(1, 4, 5, 7)
v[5]
v[10]
```


3. What do the sort, sum, mean, sd, max, min functions do?

  - `sort()` orders vectors in asc or desc order.
  - `sum()` returns the sum of all present values.
  - `mean()` returns the trimmed arithmetic mean.
  - `sd()` returns the standard deviation of values in a numeric vector or R object.
  - `max()` returns the max value in a vector.
  - `min()` returns the min value in a vector.

4. What do `which.max` and `which.min` do? **important**

  - The `which` prefix on max and min has the function return the *location* or *index* of the first min or max value of a numeric or logical vector.
  


 
