---
title: "Indiana Bloomington DSE R Notebook"
output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

## Logical, used to store TRUE or FALSE
- You can use **T** or **F** as well.

```{r}
v = TRUE
print(v)
typeof(v)
```

#### R is case sensitive.

## Numeric, used to store real values

```{r}
v <- 23.5
print(v)
typeof(v)
```

## Integer, used to store integer values

```{r}
v <- 5L
print(v)
typeof(v)
```

- Creating an integer without the **L suffix** will create it as **Numeric**.

```{r}
v = 5
typeof(v)
```

## Character, used to store characters or strings
- Both single and double quotes are acceptable.

```{r}
v <- "data"
typeof(v)
```

## Complex, used to store **complex** values

```{r}
v <- 2 + 3i
print(v)
typeof(v)
```

## Raw, used to store **raw values**; *not very important*

```{r}
v <- charToRaw("Hello")
print(v)
typeof(v)
```

# Section 2
## Atomic Vectors

### Create Vectors

```{r}
v = c(1, 2, 3, 4)
v
```

```{r}
v = c(1:100) # creates a vector with numbers 1 through 100
# v = 1:100 also works.
v
```

1. Create a vector using `v = 4.5:10.5` and `v = 4.5:11.3`:
```{r}
v = 4.5:10.5
r = 4.5:11.3

v
r
```

2. What does the function `seq()` do? (HINT: Read `seq()` docs)
- `seq()` generates a numerical sequence from the first argument to the second.

3. What does `v = c()` do? What is `NULL` ?
```{r}
v = c()
v
```

-- `NULL` is an object in R that is returned sometimes by expressions and/or functions that have a value of *undefined*.

## Vector Properties
  - `typeof()` to access data type contained in the vector;
  - `length()` to access the number of elements in the vector;
  - `attributes()` to access add'l metadata on the vector
  
##### Numeric Vector
```{r}
num_vec = c(1, 2, 3, 4)
num_vec

typeof(num_vec)
length(num_vec)
attributes(num_vec)

```

##### Character Vector
```{r}
char_vec = c("aa", "bb", "cc", "dd", "ee")
char_vec

typeof(char_vec)
length(char_vec)
attributes(char_vec)


```

##### Logical Vector

```{r}
log_vec = c(T, F, F, T)
log_vec

typeof(log_vec)
length(log_vec)
attributes(log_vec)

```

Use the following functions to query the type of a vector for a boolean result:
  - `is.atomic()` -- tells if a vector is **atomic** or not;
  - `is.numeric()` -- is the data type **Numeric** or not;
  - `is.double()` -- is the data type **Double Float** or not;
  - `is.logical()` -- is the data type **Logical** or not;
  - `is.integer()` -- is the data type **Integer** or not;
  - `is.character()` -- is the data type **Character** or not
  
## Accessing Elements of a Vector
#### Indexing Operator `[]` 
  -- Indices in **R** start with **1** **(!!)**

```{r}
v = c(1, 2, 3, 4, 5, 6, 7, 8)
v[1] # prints first element
v[c(1,2)] # prints 1st and 2nd elements
v[2:5] # prints 2nd element to the 5th
v[-1] # prints all elements except the 1st
v[-1:-3] # prints all elements except 1 to 3
v[-c(3, 1)] # prints all except 1st and 3rd
v[c(2.5, 4.6)] # real numbers are truncated to integers; prints 2nd and 4th elements
# v[c(2, -4)] # Error; cannot use negative and positive indices.
```


  