---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Matrix

  - **Matrix** : a collection of data elements arranged in a 2-D rectangular layout. Like a **vector**, but contains a *dimension* attribute.

## Create a Matrix

  - `matrix()` function creates matrices. Arguments ***nrow*** and ***ncol*** take values to define the dimensions.
  - Passing only one argument value leads the function to infer the other.

```{r}
m = matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2)
m
```

  - Use `byrow = TRUE` to fill the matrix by row, instead of column which is the default.
  
```{r}
m = matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2, byrow = TRUE)
m
```

  - Also, the functions `cbind()` and `rbind()` are available to create matrices. They take two or more vectors of the same length and bind them column-wise or row-wise respectively.
  
```{r}
x = cbind(c(1,2,3), c(4,5,6))
x

y = rbind(c(1,2,3), c(1,2,3))
y
```

## Dimension Attributes

  - `dim()` function sets dimension attribute.
  
```{r}
attributes(m) # 3 rows, 2 columns

dim(m)

dim(m)[1]

dim(m)[2] # number of columns

```

  - Change a vector into a matrix by giving it a dimension attribute using `dim()`.

```{r}
v = 1:6
dim(v) = c(2, 3) # 2 rows, 3 cols
v

is.matrix(v)
```

## Naming Matrix rows and columns

  - Pass a **2-element list** to the `dimnames` argument of `matrix()`, **row names, then col names**.
  
```{r}
m <- matrix(1:9, nrow = 3, dimnames = list(c("X", "Y", "Z"), c("A", "B", "C")))
m
```

  - Get dimension names with `dimnames()`. 
  - Access and change these names with `colnames()` and `rownames()`
  
```{r}
dimnames(m)

rownames(m)
colnames(m)

# changing row and columns names:

rownames(m) = c("X1", "Y1", "Z1")
colnames(m) = c("A1", "B1", "C1")
dimnames(m)
```

  - `attributes()` gives the names and dimensions of a matrix. Use `str()` for more detailed info:
  
```{r}
attributes(m)

str(m)
```
  
## Access to Matrix Elements

  - `[]` used thus: `var[row, column]`
  - Get multiple elements by placing a *vector* in `[]`.
  
```{r}
m[c(1, 2), c(1, 2)]

m[1:2, 1:2]

# if column field is blank, it will select all columns

m[c(1,2), ]

m[1,] # first row only

m[,1] # first column only

m[,] # whole matrix

# negative indices exclude that row/column, like vector.

m[-1,]
```

  - `drop = FALSE` passed to `[]` returns the indexed matrix as a *matrix* and not a *vector*:
  
```{r}
m[1,]
m[1, , drop = FALSE]

```
  
  - Possible to index a matrix with a vector. It stacks columns of a matrix one after the other, and returns a **vector**. 
  
```{r}
m = matrix(1:9)

m[1:4]

m[c(2,3,5,7)]
```
- Use **names** to access elements, similarly to vectors.

```{r}
m <- matrix(1:9, nrow=3, dimnames = list(c("X","Y","Z"), c("A","B","C")))
m[,"A"]

m[c("X", "Y"), c("A", "B")]
```

  - **Logical vectors** can access matrix elements. Indexing vectors are used, with logical values, like a *mask* to access values where `TRUE` is mapped.
  
```{r}
m[c(TRUE,FALSE,TRUE), c(TRUE,TRUE,FALSE)]

m[c(T,F), c(F,T,F)] # recycles as c(T,F,T), c(F,T,F)

m[c(T,F,T), c(2, 3)]

m[c(T,F,T), c("A", "C")]

```
  
## Modify a Matrix

```{r}
m = matrix(1:9, nrow=3)
m

m[2,2] = 0
m
```

  - **Transpose** a matrix with `t()`.

```{r}
t(m)
```

  - `rbind()` and `cbind()` are used here to add a row or column, or alternatively, remove them.
  
```{r}
m = cbind(m, D = c(0,0,0)) # add column D
m

```
```{r}
m = m[,1:3] # remove the fourth column
m
```
  
  - Alter the **dimensions** of a matrix:
```{r}
m = matrix(1:6, nrow = 3)
m
```
  
```{r}
dim(m) = c(2,3)
m
```
  
  