---
title: "Indiana Bloomington DSE R Notebook"
output: 
  output: 
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

# Data Frame

  - Gives us a tabular data structure.
  
## Create Data Frames

  - It is, in simple terms, a list of vectors of equal length.
  - `data.frame()` to create
  - Each vector in the list is a separate **column**.
  
```{r}
a = c(1, 2, 3, 4)
b = c(T, F, T, T)
s = c("one", "two", "three", "four")
df = data.frame(a, b, s)
df
```

```{r}
typeof(df)
```
  - Useful functions for getting info about a data frame:
  
```{r}
names(df) # names of columns

row.names(df) # names of rows

ncol(df) # number of rows

nrow(df) # number of columns

length(df) # length is same as number of columns
```
```{r}
df = data.frame(a = 1:5, b = 5:9, c = c(T,F,T,T,F), d = c("one", "two", "three", "four", "five"))
df
```
  
  - `str()` shows the structure of a data frame. "5obs" means 5 rows. "4 variables" means 4 columns. Column type and values are printed.
  
```{r}
str(df)
```

  - `data.frame()` converts strings to factors. Pass in `stringsAsFactors=FALSE` in `data.frame()` to avoid this.
  
```{r}
df = data.frame(a = 1:5, b = 5:9, c = c(T,F,T,T,F), d = c("one", "two", "three", "four", "five"), stringsAsFactors = FALSE)
str(df)
```

  - `[]` **preserves** data type; output is same as input.
  - `[[]]` **simplifies** data type, returns the simplest possible structure that can represent the output.
  
```{r}
df

df["a"]


```

```{r}
df[1]
```

```{r}
df$a
```

```{r}
df[["a"]]
```

```{r}
df[[1]]
```

  - Data Frames can be accessed with row and column indexing:

```{r}
df[1, 2] # gives element at 1st row and 2nd column index

df[2, ] # gives the entire 2nd row

df[, 1] # gives the 1st column

# Use the drop = FALSE argument tot he index operator to return a dataframe rather than a vector.

df[, 1, drop = FALSE]
```

  - Vectors of indices work too.
```{r}
df[1:2,2:3]
```
  
  - **Negative Indexing**
```{r}
df[-1, ]
```
  
## Modify a Data Frame

  - Modify through **reassignment**, like vectors and matrices.

```{r}
df[1,2] = 10
```

  - Add rows or cols with `rbind()` or `cbind()`:
```{r}
df = rbind(df, list(11, 12, FALSE, "six"))
df
```
  
```{r}
df = cbind(df, e = 15:20)
df
```

## Delete Elements

  - Assign a column the `NULL` value to delete them.
  
```{r}
df$e = NULL
df
```
  
  - Also, **reassignment** will delete rows.
```{r}
df = df[-1, ]
df
```
  
## Build-in Data Frame

  - `mtcars` and `iris` are built-in data frames to use.
  
```{r}
str(mtcars)
```

```{r}
str(iris)
```

  - Use `head` to see the first few lines of a data frame. 
  - Use `tail` to see the last few lines.
  
```{r}
head(mtcars)
```

```{r}
tail(mtcars)
```

